DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car

### 2017-10-16 paper/appendix.tex
Changes to paper

In the first method, we trained the neural network model across a set 
of 30 completed runs on the track seen in Figure~\ref{fig:track} by a
human pilot. Half of the runs saw the car driving one way along the
track, while the remaining half were of the car driving in the
opposite direction on the track.
In total, we collected 2,556 frames for training and 2,609 
frames for validation.
The weights of the network are initialized using Xavier
initializer~\cite{Glorot2010}, which is known to provide better
initial values than the random weight assignment method.
In each training step, we use a batch
size of 100 frames, which are randomly selected among all the
collected training images, to optimize the network.
We repeat this across 2,000 training steps. When a model was trained
with the  aformentioned data, the training loss was 0.0188 and the
validation  loss was 0.0132.
The change of the loss value over the course of model training can be
seen in Figure~\ref{fig:modelloss}.

In the second method, we use the same data and parameters as  above 
except that now images are labeled as 'curved' and 'straight' and we pick an
equal number of images from each category at each training step to
update the model. In other words, we try to remove bias in selecting
images. We find that the car performed better in practice by applying
this approach as the car displayed a greater ability to stay in the
center of the track (on the white tape).
However, we find that there is a huge discrepency between the training
loss and the validation loss as the former was 0.009, while the latter
was 0.0869---a 10X difference---indicating that the model suffers from
the overfitting problem.

@@

%% We believe, though, that this can be
%% rectified through the collection of additional data for training and
%% validation.

%% \begin{figure*}[h]
%%   \centering
%%   \includegraphics[width=.9\textwidth]{figs/graph_run=.pdf}
%%   \caption{ Network visualization. }
%%   \label{fig:}
%% \end{figure*}
%% \fixme{Is all of the network graph necessary or only certain parts?}

@@

%% Specifically, these issues all have the potential to negatively
%% affect the cpu clock speed/frequency, which would result in decreased
%% performance. 
In all our experiments on the Raspberry Pi 3, the CPU  operated at a
preferred clock speed of 1.2 GHz. However, without care, it is
possible for the CPU may be operating at a lower frequency.

An important factor is CPU thermal throttling, which can affect CPU
clock speed if the CPU temperature is too high (Pi 3's firmware is
configured to throttle at 85C).
DNN model operations are computationally intensive, thus it is
possible for the temperature of the CPU to rise quickly. This can be
especially problematic in situations where multiple DNN models are
running simultaneously on the Pi 3.
If the temperature reaches the threshold, the Pi 3's thermal
throttling kicks in and decreases the clock speed down to 600MHz---a
half of the maximum 1.2GHz---so that the CPU's temperature stays at a
safe level.
We found that without proper cooling solutions (heatsink or fan), 
prolonged use of the system would result in CPU frequency decrease
that may affect evaluation.

@@

%% Thus the
%% Raspberry Pi may not be suited for prolonged use, especially in cases
%% where the workload is relatively larger, such as running multiple
%% models. Rather, the Pi seems to be better suited for running in set
%% periods, after which it is turned  off or made idle so that the cpu is
%% allowed time to cool down.

>>>>

In the first method, we trained the neural network model across a set 
of 30 completed runs on the track seen in Figure~\ref{fig:track} by a
human pilot. Half of the runs saw the car driving one way along the
track, while the remaining half were of the car driving in the
opposite direction on the track.
In total, we collected 2,556 frames for training and 2,609 
frames for validation.
The weights of the network are initialized using a Xavier
initializer~\cite{Glorot2010}, which is known to provide better
initial values than the random weight assignment method.
In each training step, we use a batch
size of 100 frames, which are randomly selected among all the
collected training images, to optimize the network.
We repeat this across 2,000 training steps. When a model was trained
with the  aformentioned data, the training loss was 0.0188 and the
validation  loss was 0.0132.
The change of the loss value over the course of model training can be
seen in Figure~\ref{fig:modelloss}.

In the second method, we use the same data and parameters as  above 
except that now images are labeled as 'curved' and 'straight' and we pick an
equal number of images from each category at each training step to
update the model. In other words, we try to remove bias in selecting
images. We find that the car performed better in practice by applying
this approach as the car displayed a greater ability to stay in the
center of the track (on the white tape).
However, we find that there is a huge discrepency between the training
loss and the validation loss as the former was 0.009, while the latter
was 0.0869---a 10X difference---indicating that the model suffers from
an overfitting problem.

@@

@@

In all our experiments on the Raspberry Pi 3, the CPU  operated at a
preferred clock speed of 1.2 GHz. However, without care, it is
possible for the CPU to operate at a lower frequency.

An important factor is CPU thermal throttling, which can affect CPU
clock speed if the CPU temperature is too high (Pi 3's firmware is
configured to throttle at 85C).
DNN model operations are computationally intensive, thus it is
possible for the temperature of the CPU to rise quickly. This can be
especially problematic in situations where multiple DNN models are
running simultaneously on the Pi 3.
If the temperature reaches the threshold, the Pi 3's thermal
throttling kicks in and decreases the clock speed down to 600MHz---
half of the maximum 1.2GHz---so that the CPU's temperature stays at a
safe level.
We found that without proper cooling solutions (heatsink or fan), 
prolonged use of the system would result in CPU frequency decrease
that may affect evaluation.

@@

### 2017-10-16 paper/background.tex
Changes to paper

The use of neural networks for end-to-end control of autonomous
car was first demonstrated in late 1980s~\cite{Pomerleau1989},
using a small 3-layer fully connected neural network; and subsequently
in a DARPA Autonomous Vehicle (DAVE) project in early
2000s~\cite{LeCun:04}, using a 6 layer convolutional neural network
(CNN); and most recently in NVIDIA's DAVE-2
project~\cite{Bojarski2016}, using a 9 layer CNN. In all of these projects,
the neural network models take raw image pixels as input and directly
produce steering control commands, bypassing all intermediary steps and
hand-written rules used in conventional robotics control approach.  
NVIDIA's latest effort reports that their trained CNN
autonomously controls their modified cars on public roads without human
intervention~\cite{Bojarski2016}.

Using deep neural networks involves two distinct
phases~\cite{NVIDIA2015}. The first
phase is \emph{training} during which the weights of the network are
incrementally updated by backpropagating errors it sees from the
training examples. Once the network is trained---i.e., the weights of
the network minimize errors in the training examples---the next phase
is \emph{inferencing}, during which unseen data is fed to the network
as input to produce predicted output (e.g., predicted image
classification). In general, the training phase is most computationally
intensive and requires high throughput, which is generally not
available on embedded platforms. The inferencing phase, on the
other hand, is relatively less computationally intensive and latency becomes
as important, if not moreso, as computational throughput, because many
use-cases have strict real-time requirements (e.g., search query
latency).

>>>>

The use of neural networks for end-to-end control of autonomous
car was first demonstrated in late 1980s~\cite{Pomerleau1989},
using a small 3-layer fully connected neural network; and subsequently
in a DARPA Autonomous Vehicle (DAVE) project in early
2000s~\cite{LeCun:04}, using a 6 layer convolutional neural network
(CNN); and most recently in NVIDIA's DAVE-2
project~\cite{Bojarski2016}, using a 9 layer CNN. In all of these projects,
the neural network models take raw image pixels as input and directly
produce steering control commands, bypassing all intermediary steps and
hand-written rules used in the conventional robotics control approach.  
NVIDIA's latest effort reports that their trained CNN
autonomously controls their modified cars on public roads without human
intervention~\cite{Bojarski2016}.

Using deep neural networks involves two distinct
phases~\cite{NVIDIA2015}. The first
phase is \emph{training} during which the weights of the network are
incrementally updated by backpropagating errors it sees from the
training examples. Once the network is trained---i.e., the weights of
the network minimize errors in the training examples---the next phase
is \emph{inferencing}, during which unseen data is fed to the network
as input to produce predicted output (e.g., predicted image
classification). In general, the training phase is more computationally
intensive and requires high throughput, which is generally not
available on embedded platforms. The inferencing phase, on the
other hand, is relatively less computationally intensive and latency becomes
as important, if not moreso, as computational throughput, because many
use-cases have strict real-time requirements (e.g., search query
latency).

### 2017-10-16 paper/evaluation.tex
Changes to paper

For real-time control of a car (or any robots), the control loop
frequency must be sufficiently high so that the car can quickly
react to the changing environment and its internal states. In general,
control performance improves when the frequency is higher, though
computation time and the type of particular physical system are
factors in determining a proper control loop frequency. While a standard
control system may be comprised of multiple control loops with
differing control frequencies---e.g., a inner control loop for lower-level
PD control, a outer loop for motion planning, etc.---DeepPicar's
control loop is a single layer, as shown earlier in
Figure~\ref{fig:controlloop}, since a single deep neural network
replaces the traditional multi-layer control pipline. (Recall
Figure~\ref{fig:end-to-end-control} on the differences between the
standard robotics control vs. end-to-end deep learning approach).
This means that the DNN interference operation must be completed
within the inner-most control loop update frequency. To understand
achievable control-loop update frequencies, we experimentally measured
the execution times of DeepPicar's DNN inference operations.

@@

Figure~\ref{fig:control-loop-timing} shows the measured control loop 
processing times of the DeepPicar over 1000 image frames (one per each
control loop). We omit the first frame processing time for cache
warmup. Table~\ref{tbl:control-loop-breakdown} shows the time
breakdown of each control loop. Note that all four CPU cores of the
Raspberry Pi 3 were used by the TensorFlow library when performing the
DNN inference operations.

First, as expected, we find that the inference operation
dominates the control loop execution time, accounting for about 85\% of
the execution time.
Second, and more importantly, we also find that the measured average
execution time of a single control loop is 42.67 ms, or 23.4 Hz and
the 99 percentail time is 50.70 ms.
This means that the DeepPicar can operate
at about 20 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, without needing any remote computing
resources. We consider these results respectable given the complexity
of the deep neural network, and the fact that the inference operation
performed by TensorFlow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow.)

In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, reportedly run at 30 Hz~\cite{Bojarski2016},
just a bit faster than the DeepPicar. Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TX2, later in
Section~\ref{sec:comparison}), the fact that the low-cost
Raspberry Pi 3 can achieve similar real-time control performance is
surprising.

>>>>

For real-time control of a car (or any robot), the control loop
frequency must be sufficiently high so that the car can quickly
react to the changing environment and its internal states. In general,
control performance improves when the frequency is higher, though
computation time and the type of the particular physical system are
factors in determining a proper control loop frequency. While a standard
control system may be comprised of multiple control loops with
differing control frequencies---e.g., an inner control loop for lower-level
PD control, an outer loop for motion planning, etc.---DeepPicar's
control loop is a single layer, as shown earlier in
Figure~\ref{fig:controlloop}, since a single deep neural network
replaces the traditional multi-layer control pipline. (Refer to
Figure~\ref{fig:end-to-end-control} on the differences between the
standard robotics control vs. end-to-end deep learning approach).
This means that the DNN inference operation must be completed
within the inner-most control loop update frequency. To understand
achievable control-loop update frequencies, we experimentally measured
the execution times of DeepPicar's DNN inference operations.

@@

Figure~\ref{fig:control-loop-timing} shows the measured control loop 
processing times of the DeepPicar over 1000 image frames (one per each
control loop). We omit the first frame's processing time for cache
warmup. Table~\ref{tbl:control-loop-breakdown} shows the time
breakdown of each control loop. Note that all four CPU cores of the
Raspberry Pi 3 were used by the TensorFlow library when performing the
DNN inference operations.

First, as expected, we find that the inference operation
dominates the control loop execution time, accounting for about 85\% of
the execution time.
Second, and more importantly, we also find that the measured average
execution time of a single control loop is 42.67 ms, or 23.4 Hz and
the 99 percentile time is 50.70 ms.
This means that the DeepPicar can operate
at about a 20 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, as no remote computing resources 
were necessary. We consider these results respectable given the complexity
of the deep neural network, and the fact that the inference operation
performed by TensorFlow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow).

In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016},
just a bit faster than the DeepPicar. Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TX2, later in
Section~\ref{sec:comparison}), the fact that the low-cost
Raspberry Pi 3 can achieve similar real-time control performance is
surprising.

### 2017-10-16 paper/intro.tex
Changes to paper

% related work and remaining problems
To understand what kind of computing hardware is needed for AI
workloads, we need a testbed and realistic workloads. While using a real 
car-based testbed would be most ideal, it is not only highly expensive, but also
poses serious safety concerns that hinder development and exploration.
Therefore, there is a strong need for safer and less costly
testbeds. There are already several relatively inexpensive RC-car
based testbeds, such as MIT 
RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
However, these RC-car testbeds still cost more than \$3,000, requiring
considerable investment.

@@

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
platform for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs state-of-the-art AI
technologies, including a vision-based end-to-end control system that
utilizes a deep convolutional neural network (CNN).
The network receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The network has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
The network architecture is identical to that of NVIDIA's DAVE-2
self-driving car~\cite{Bojarski2016}, which uses a much more powerful
computer (Drive PX computer~\cite{drivepx}) than Raspberry Pi~3.
We chose to use Pi 3 not only because of its
affordability, but also because it is representative
of today's mainstream low-end embedded multicore platforms found in
smartphones and other embedded devices.

@@

% how we trained (maybe moved to a later section)
We apply a standard imitation learning methodology to train the car to
follow tracks on the ground. We collect data for
training and validation by manually
controlling the RC car and recording the vision (from the webcam
mounted on the RC-car) and the human control inputs. We then train the
network offline using the collected data on a desktop computer, which
is equipped with a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
back to the Raspberry Pi, which is then used to perform inference
operations---locally on the Pi---in the car's main control loop in
real-time. For real-time control, each inference operation must
be completed within the desired control period (e.g., 50ms period for
20Hz control frequency).
% how we evaluated (in terms real-time performance)

% what are our findings?
Using the DeepPicar platform, we systematically analyzed its real-time
capabilities in the context of deep-learning based real-time
control, especially on real-time deep neural network inferencing.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's deep-learning based control system and the impact of
computing hardware architectures.

@@

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present the design and implementation of a
    low-cost autonomous vehicle testbed DeepPicar, which utilizes
    state-of-the-art artificial intelligence techniques.
  \item We provide an analysis and case-study of real-time issues in the
    DeepPicar.
  \item We systematically compare real-time computing capabilities of
    multiple embedded computing platforms in the context of
    vision-based autonomous driving.
\end{itemize}

>>>>

% related work and remaining problems
To understand what kind of computing hardware is needed for AI
workloads, we need a testbed and realistic workloads. While using a real 
car-based testbed would be most ideal, it is not only highly expensive, but also
poses serious safety concerns that hinder development and exploration.
Therefore, there is a strong need for safer and less costly
testbeds. There are already several relatively inexpensive RC-car
based testbeds, such as MIT's 
RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
However, these RC-car testbeds still cost more than \$3,000, requiring
considerable investment.

@@

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
platform for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs state-of-the-art AI
technologies, including a vision-based end-to-end control system that
utilizes a deep convolutional neural network (CNN).
The network receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The network has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
The network architecture is identical to that of NVIDIA's DAVE-2
self-driving car~\cite{Bojarski2016}, which uses a much more powerful
computer (Drive PX computer~\cite{drivepx}) than a Raspberry Pi~3.
We chose to use a Pi 3 not only because of its
affordability, but also because it is representative
of today's mainstream low-end embedded multicore platforms found in
smartphones and other embedded devices.

@@

% how we trained (maybe moved to a later section)
We apply a standard imitation learning methodology to train the car to
follow tracks on the ground. We collect data for
training and validation by manually
controlling the RC car and recording the vision (from the webcam
mounted on the RC-car) and the human control inputs. We then train the
network offline using the collected data on a desktop computer, which
is equipped with a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
back to the Raspberry Pi, which is then used to perform inference
operations---locally on the Pi---in the car's main control loop in
real-time. For real-time control, each inference operation must
be completed within the desired control period (e.g., 50 ms period for
20 Hz control frequency).
% how we evaluated (in terms real-time performance)

% what are our findings?
Using the DeepPicar platform, we systematically analyze its real-time
capabilities in the context of deep-learning based real-time
control, especially on real-time deep neural network inferencing.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's deep-learning based control system and the impact of
computing hardware architectures.

@@

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present the design and implementation of a
    low-cost autonomous vehicle testbed, DeepPicar, which utilizes
    state-of-the-art artificial intelligence techniques.
  \item We provide an analysis and case-study of real-time issues in the
    DeepPicar.
  \item We systematically compare real-time computing capabilities of
    multiple embedded computing platforms in the context of
    vision-based autonomous driving.
\end{itemize}

### 2017-10-16 paper/main.tex
Changes to paper

\maketitle
\thispagestyle{empty}
\begin{abstract}
We present DeepPicar, a low-cost deep neural network (DNN) based
autonomous car platform. DeepPicar is a small scale
replication of a real self-driving car called Dave-2 by NVIDIA, which
drove on public roads using a deep convolutional neural network (CNN), 
that takes images from a front-facing camera as input and produces
car steering angles as output. DeepPicar uses the exact same network
architecture---9 layers, 27 million connections and 24K
parameters---and can be trained to drive itself, in real-time, using a
web camera and a modest Raspberry Pi 3 quad-core platform.
Using DeepPicar, we analyze Pi 3's computing capabilities to 
support end-to-end deep learning based real-time control of autonomous
vehicles. We also systematically compare other contemporary embedded
computing platforms using the DeepPicar's CNN based real-time control
software as a workload. 
We find all tested platforms, including the Pi 3, are capable of
supporting deep-learning based real-time control, from 20 Hz up to 100
Hz depending on hardware platform. 
However, shared resource contention remains an
important issue that must be considered in applying deep-learning
models on shared memory based embedded computing platforms.
\end{abstract}

>>>>

\maketitle
\thispagestyle{empty}
\begin{abstract}
We present DeepPicar, a low-cost deep neural network (DNN) based
autonomous car platform. DeepPicar is a small scale
replication of a real self-driving car called Dave-2 by NVIDIA, which
drove on public roads using a deep convolutional neural network (CNN), 
that takes images from a front-facing camera as input and produces
car steering angles as output. DeepPicar uses the exact same network
architecture---9 layers, 27 million connections and 250K
parameters---and can be trained to drive itself, in real-time, using a
web camera and a modest Raspberry Pi 3 quad-core platform.
Using DeepPicar, we analyze the Pi 3's computing capabilities to 
support end-to-end deep learning based real-time control of autonomous
vehicles. We also systematically compare other contemporary embedded
computing platforms using the DeepPicar's CNN based real-time control
software as a workload. 
We find all tested platforms, including the Pi 3, are capable of
supporting deep-learning based real-time control, from 20 Hz up to 100
Hz depending on hardware platform. 
However, shared resource contention remains an
important issue that must be considered in applying deep-learning
models on shared memory based embedded computing platforms.
\end{abstract}

### 2017-10-16 paper/overview.tex
Changes to paper

In this section, we provide an overview of our DeepPicar platform. In
developing DeepPicar, one of our primary goals is to replicate the
NVIDIA's DAVE-2 system on a smaller scale with using a low cost
multicore platform, Raspberry Pi 3. Because Pi 3's computing
performance is much lower than that of the DRIVE PX platform used in
DAVE-2, we are interested in if, and how, we can process
computationally expensive neural network operations in
real-time. Specifically, inferencing (forward pass processing)
operations must be completed within each control period
duration---e.g., WCET of 50ms for 20Hz control frequency---locally on
the Pi 3 platform, although training of the network (back-propagation
for weight updates) can be done offline and remotely using a desktop
computer.

@@

% inferencing on pi3
Once the network is trained on the desktop computer, the trained model
is copied back to the Raspberry Pi. The network is then used
by the car's main controller, which feeds a image frame from the web
camera as input to the network. The produced steering angle output is
then converted as the PWM values of the steering motor of the
car. Figure~\ref{fig:controlloop} shows simplified pseudo code of the
controller's main loop. Among the five steps, step 3, network
inferencing, is the most computationally intensive and dominates the
execution time.

@@

Note that although the steering angle output of the network $angle$ is
a continuous real value, the RC car platform we used unfortunately
only supports three discrete angles---left (-30$^{\circ}$), center
(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
Currently, we approximate the network generated real-valued
angle to the closest one of the three angles, which may
introduce inaccuracy in control.
In the future, we plan to use a different (more expensive) RC car
platforms that can precisely control the car's steering angle.
We would like to stress, however, that the use of different RC car
platforms has no impact in terms of computational 
aspects of the system, and that our main focus of this study is
not in improving network accuracy but in closely replicating the
DAVE-2's network architecture and its real-time characteristics.

Another issue we observed in training/testing the network, which could
affect the network performance, is camera latency. In DeepPicar's
context, the camera latency is from the time the camera
sensor observes the scene to the time the computer actually reads the
digitized image data. Unfortunately, this time can be significantly
long depending on camera and the performance of the Pi. We
experimentally measure DeepPicar's camera latency as
follows~\footnote{We follow the method and the tool found at
  https://www.makehardware.com/2016/03/29/finding-a-low-latency-webcam/}.
First, we point the DeepPicar's camera toward to a computer monitor, which is
conntected to the Raspberry Pi 3's HDMI output. We then launch a 
program that overlays the current time with the camera's preview
screen on the monitor. Then, an infinite number of time strings will be
shown on the monitor screen and the time difference between two
successive time strings represent the camera
latency. Figure~\ref{fig:camera-latency} shows the captured moniotor
screen. The measured camera latency is about 300-350 ms. This is
significantly higher than the latency of human perception, which is known
to be as fast as 13 ms~\cite{ThomasBurger2015}. Higher camera latency
could negatively affect to the control performance, because the DNN would
analyze stale scenes. In the future, we plan to identify and use
low-latency cameras.

>>>>

In this section, we provide an overview of our DeepPicar platform. In
developing DeepPicar, one of our primary goals is to replicate the
NVIDIA's DAVE-2 system on a smaller scale with using a low cost
multicore platform, Raspberry Pi 3. Because the Pi 3's computing
performance is much lower than that of the DRIVE PX platform used in
DAVE-2, we are interested in if, and how, we can process
computationally expensive neural network operations in
real-time. Specifically, inferencing (forward pass processing)
operations must be completed within each control period
duration---e.g., a WCET of 50ms for 20Hz control frequency---locally on
the Pi 3 platform, although training of the network (back-propagation
for weight updates) can be done offline and remotely using a desktop
computer.

@@

% inferencing on pi3
Once the network is trained on the desktop computer, the trained model
is copied back to the Raspberry Pi. The network is then used
by the car's main controller, which feeds a image frame from the web
camera as input to the network. In each control period, the produced 
steering angle output is then converted as the PWM values of the 
steering motor of the car. Figure~\ref{fig:controlloop} shows simplified 
pseudo code of the controller's main loop. Among the five steps, the 3rd step, 
network inferencing, is the most computationally intensive and dominates the
execution time.

@@

Note that although the steering angle output of the network $angle$ is
a continuous real value, the RC car platform we used unfortunately
only supports three discrete angles---left (-30$^{\circ}$), center
(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
Currently, we approximate the network generated real-valued
angle to the closest one of the three angles, which may
introduce inaccuracy in control.
In the future, we plan to use a different (more expensive) RC car
platform that can precisely control the car's steering angle.
We would like to stress, however, that the use of different RC car
platforms has no impact on the computational 
aspects of the system, and that our main focus of this study is
not in improving network accuracy but in closely replicating the
DAVE-2's network architecture and its real-time characteristics.

Another issue we observed in training/testing the network, which could
affect the network performance, is camera latency. In DeepPicar's
context, the camera latency is from the time the camera
sensor observes the scene to the time the computer actually reads the
digitized image data. Unfortunately, this time can be significantly
long depending on the camera and the performance of the Pi. We
experimentally measure DeepPicar's camera latency as
follows~\footnote{We follow the method and the tool found at
  https://www.makehardware.com/2016/03/29/finding-a-low-latency-webcam/}.
First, we point the DeepPicar's camera towards a computer monitor, which is
conntected to the Raspberry Pi 3's HDMI output. We then launch a 
program that overlays the current time with the camera's preview
screen on the monitor. Then, an infinite number of time strings will be
shown on the monitor screen and the time difference between two
successive time strings represent the camera
latency. Figure~\ref{fig:camera-latency} shows the captured moniotor
screen. The measured camera latency is about 300-350 ms. This is
significantly higher than the latency of human perception, which is known
to be as fast as 13 ms~\cite{ThomasBurger2015}. Higher camera latency
could negatively affect to the control performance, because the DNN would
analyze stale scenes. In the future, we plan to identify and use
low-latency cameras.
